---
title: "DATA ANALYTICS WITH R - PROJECT"
author: "Nikitha Jayakumar, Sakshi Shetty"
date: "3/1/2019"
output: html_document
---

For this project, we shall be building a model to predict the values of the attribute suicide number.

```{r}
dataset <- read.csv("C:\\Users\\Nikitha\\Documents\\PESU IO\\Project\\who_suicide_statistics.csv")
```

The dataset is chosen such that suicide statistics is considered for the years after 2000 and of those countries where population is given.
Hence, following will be the dataset:

```{r}
data <- subset(dataset,year>=2000 & is.na(population)==FALSE)
```

Preprocessing the data:

Filling in the missing values with mean.

```{r}
data$suicides_no[is.na(data$suicides_no)] <- mean(data$suicides_no, na.rm=TRUE)   
```

Finding the correlation between numeric attributes:

```{r}
sapply(data, class)
print(cor(data[c("year","suicides_no","population")]))
```

As can be seen by the correlation matrix, the correlation is highest between suicides_no and population; leading to the inference that there are a higher number of suicides in larger populations.
Thus, the attribute population must be included in the regression model.


To determine the relations between categorical variables and the target:

By visulaising the scatter plot, 
1. It becomes evident that the suicide rate is highest for the age group 35-45       years.
2. Rate of suicide is higher in males than females.
3. Starting from the year 2000, suicides seemed to have peaked between the years     2000 and 2001 and then show a gradual decrease to an almost constant level        between 2010-2015.

From the plots, it can be seen that the attributes of age, sex, country and year will have to be considered for the regression model.

```{r}
library(plotly)
print(plot_ly(data,x=~age,y=~suicides_no,trace="scatter"))
print(plot_ly(data,x=~sex,y=~suicides_no,trace="scatter"))
print(plot_ly(data,x=~year,y=~suicides_no,trace="scatter",mode="lines"))
print(plot_ly(data,x=~country,y=~suicides_no,trace="scatter"))
```



To split the data into training and testing sets using 80:20 split.

```{r}
train_index <- sample(floor(nrow(data)*0.8))
train<- data[train_index,]
test <- data[-train_index,]
```

To build the regression model.

```{r}
model <- lm(suicides_no~sex+population+year+country+age,train)
```

To predict values using the model built.

```{r}
pred <- predict(model,data=test)
```

To calculate coefficient of determination and plot graphs for the model.

```{r}
summary(model)
plot(model)
plot(residuals(model))
```

